Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated

### 2017-01-01 ms_pre_post.Rmd
Remove comments; finalise version to send to coauthors

<!--- Will most likely need to read and cite \cite{Fisher2016} here (Fisher et al. 2016 Behavioural Ecology -- pre- and post-copulatory competition using network theory --->

@@

<!--- I have removed 'mate' in front of 'choice' and replaced it with 'reproductive decisions' in the second sentence because post-copulatory inbreeding avoidance isn't an expression of mate choice; mating has already happened at this point; females are choosing paternity among the mates that they already chose --->

<!--- Note that there are twice as many butterfly species and about eight times as many weevil species as there are bird species. I think that weevils and butterflies can be on at least equal footing with birds in terms of generality --->

@@

<!--- This is getting awfully narrow in terms of theory; postcopulatory *might* happen given a fairly specific set of conditions concerning evolution of pre-copulatory inbreeding avoidance and polyandry. Are we confident these conditions are general? --->

<!--- I'm not suggesting that things need to be changed (probably too late for that), just that, to me, this reads as the theoretical equivalent of a very system-specific question being address, not like a question that has broad implications across different systems --->

@@

<!--- This last sentence sounds good and strong, but it doesn’t actually follow logically from the rest of the paragraph – how do we know that we can’t predict or understand evolutionary outcomes with non-dynamic models, perhaps one that predicts long-term equilibria? This doesn’t seem obvious or well argued if we’re just talking about costs and benefits – what, specifically, can a dynamic model (I assume, a simulation, i.e., a model where things change over time) address that another type of model cannot? --->

@@

The probability that an offspring of a focal female $i$ is sired by any one of $i$'s mates $j$ is calculated by assigning a perceived fertilisation quality to each $j$ with respect to $i$, $Q^{f}_{ij}$. Perceived fertilisation quality $Q^{f}_{ij}$ is calculated in the same way as perceived mate quality $Q^{m}_{ij}$, such that if a focal female has a strategy of post-copulatory inbreeding avoidance ($F_{p}<0$), then the perceived quality of male $j$ is $Q^{f}_{ij} = (-F_{p} \times k_{ij} + 1)^{-1}$. If the focal female has strategy of post-copulatory inbreeding preference ($F_{p}>0$), then the perceived quality of male $j$ is $Q^{f}_{ij} = F_{p} \times k_{ij} + 1$. A relative quality ($q^{f}_{ij}$) is then calculated for each male by dividing his $Q^{f}_{ij}$ by the sum of the $Q^{f}_{ij}$ values across all of a female's mates. These $q^{f}_{ij}$ values, which lie between zero and one, define the probability of paternity. As with mate choice, $q^{f}_{ij}$ values are placed into a vector whose elements sum to unity, and a random uniform number is used to sample paternity. Females produce $n$ offspring, so $q^{f}_{ij}$ values are sampled $n$ times independently and with replacement for each female to determine the realised distribution of sires. <!--- This last sentence has now been added back in, with another explaining it further. These things seem to get taken out as extraneous, but then added back in because the methods are not clear when they are removed --->

@@

<!--- Note given previous comments: I've reworded because we don't know that pre-copulatory inbreeding avoidance is facilitating evolution of polyandry, not anymore. With the addition of post-copulatory strategy, it might be hindering it instead. If we wanted the answer to this, we would need to fix different levels of post-copulatory inbreeding avoidance and see if we get the same pattern as in Duthie et al. (2016) Evolution. We might not see the same result of pre-copulatory inbreeding avoidance facilitating evolution of polyandry. --->

<!--- The above point is especially problematic here because in the absence of post-copulatory inbreeding avoidance (in the current model) pre-copulatory inbreeding avoidance definitely does not facilitate evolution of costly polyandry -- this only happens under the severe constraints of our previous Evolution paper. Hence, it would have been a bit strange to suggest that post-copulatory inbreeding avoidance works `alongside' pre-copulatory inbreeding avoidance when we know that pre-copulatory inbreeding avoidance isn't doing anything. --->

@@

<!--- I am happy to switch back to using statements as titles instead of questions --->

@@

<!--- Changed several things back here because alleles cannot evolve --->

@@

However, when both pre-copulatory and post-copulatory inbreeding avoidance could evolve, their relative evolutionary dynamics depended on their relative costs. When one inbreeding strategy had zero cost, the other strategy initially evolved towards inbreeding avoidance (i.e., negative $M_{a}$ or $F_{a}$) but then evolved back towards random mating or fertilisation (i.e., $M_{a}=0$ or $F_{a}=0$) across generations following increasing evolution of inbreeding avoidance through the cost-free strategy (Figure 3B,D). <!--- For example, given fixed $F_{a}=0$, grand mean $M_{a}$ value in the last $10000$ generations was $-0.868$ (Figure 3A), meaning that a female was ca $5.34$ times less likely to choose a full-sibling and ca $2.10$ times less likely to mate with a first cousin than a non-relative given a high cost of pre-copulatory inbreeding avoidance, $c_{M}=0.02$. However, when $F_{a}$ allele values could also evolve, grand mean $M_{a}$ allele value in the last $10000$ generations was only $-0.288$ (Figure 3B). Further, given fixed $M_{a}=0$, grand mean $F_{a}$ value in the last $10000$ generations was $-1.092$ (Figure 3C), meaning that a female was ca $6.46$ times less likely to choose a full-sibling and ca $2.37$ times less likely to choose a first cousin than a non-relative given a high cost of pre-copulatory inbreeding avoidance, $c_{F}=0.02$. However, when $M_{a}$ values could also evolve and $P_{a}$ values were negative, grand mean $F_{a}$ value in the last $10000$ generations was only $-0.070$ (Figure 3D). --->

Evolution of post-copulatory inbreeding avoidance facilitated evolution of polyandry (Figure 3B,C), while monandry evolved in the absence of post-copulatory inbreeding avoidance (Figure 3A,D; polyandry was cost free in all these simulations). <!--- Overall, evolution of a costly inbreeding strategy that readily evolved in isolation was greatly impeded by evolution of a less costly strategy. --->

@@

<!--- Get in the idea that costs could interact with benefits of degeneracy in a similar way that mutation rate and gene function interact? -- Nowak --->

@@

<!--- Persistence of degenerate inbreeding strategies in spite of cost asymmetries can be interpreted as analagous to the persistence of degenerate alleles across multiple loci when genotype fitness and mutation rates vary [@Nowak1997]. --->

<!--- Another type of cost relevant for polyandry and inbreeding strategy is risk of male harm caused by sexual conflict [e.g., @Arnqvist2005a; @Parker2006]. We assumed that males were passive in mating encounters with females, but inbreeding theory predicts that males should benefit from a higher tolerance of inbreeding than females, leading to sexual conflict over inbreeding [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. Females that express the choice to avoid inbreeding when mating might therefore pay a cost of harm caused by coercion expressed by related males. This cost might itself evolve over time as male harm coevolves with female resistance [@Ronn2007; @Perry2011; @Wang2015a], and costs might vary with interactions among male non-relatives and relatives because effects of kin selection are predicted to reduce harm caused by related male suitors [@Rankin2011a]. To predict evolution of polyandry and inbreeding strategy in populations characterised by active males and sexual conflict, consideration of male inbreeding strategy and coevolution between male harm and female resistance might therefore need to be considered explicitly. --->

@@

<!--- Some general conclusions addressing 1) what future theory should be developed, and 2) what future empirical studies need to consider. For 2, need to be aware of both pre- and post-copulatory inbreeding avoidance, potentially, and that the two are not independent, necessarily; cannot find any general theory for what happens when you have this kind of strategy redundancy -- two ways of getting at the same outcome. See: @Nowak1997, @Edelman2001, @Whitacre2010, @Price2002, @Park2013, Atamas2009, @Whitacre2012, @Whitacre2010a, @Levin2008 --->

<!--- LEFT OFF HERE ------------------------------------------------------->

@@

<!--- Similar comment to above - I’m finding the structure of this paragraph a bit backwards. Might it be more helpful to give the conclusion drawn from our model and then explain how it fits with existing empirical studies? At the moment I read the stuff about the empirical studies without having any idea how it was going to be linked into our results – so then I had to go back and read it again once you’d told me the link at the end. I’m still not quite sure I get it, or why we need so much detail on mechanisms? Also, I think the writing could be greatly shortened here - could probably say more or less the same thing in half the words. I’ve done some editing to illustrate this - but I doubt this’ll be a definitive version because I can’t really see where this para is going. -->

<!--- Regarding last sentence: ''OK this is a much better opening to this section – states the key result from our model and then goes on the develop/consider the implications. Could you use the same kind of structure for the previous sections?'' --->

<!--- Our model assumed that allele values underlying polyandry, pre-copulatory inbreeding avoidance, and post-copulatory inbreeding avoidance were all initialised at zero (causing monandry, and random mating and fertilisation with respect to kinship), and phenotypes evolved simultaneously after initialisation. However, these initial conditions are likely to be atypical, as evolution of pre-copulatory and post-copulatory inbreeding avoidance from ancestral random mating and fertilisation is unlikely to occur simultaneously. Evolution of a single inbreeding strategy (e.g., pre-copulatory inbreeding avoidance) might occur in the absence of its complementary strategy (e.g., post-copulatory inbreeding avoidance) if the complementary strategy is evolutionarily constrained. Alternatively, initial evolution of one inbreeding strategy might occur in the context of an already established adaptive strategy for avoiding inbreeding. We showed that selection for post-copulatory but not pre-copulatory inbreeding avoidance was strengthened if its complementary strategy was fixed for random choice with respect to kinship, and that selection for post-copulatory but not pre-copulatory inbreeding avoidance was precluded if its complementary strategy was fixed for adaptive inbreeding avoidance. --->

>>>>

@@

@@

@@

@@

The probability that an offspring of a focal female $i$ is sired by any one of $i$'s mates $j$ is calculated by assigning a perceived fertilisation quality to each $j$ with respect to $i$, $Q^{f}_{ij}$. Perceived fertilisation quality $Q^{f}_{ij}$ is calculated in the same way as perceived mate quality $Q^{m}_{ij}$, such that if a focal female has a strategy of post-copulatory inbreeding avoidance ($F_{p}<0$), then the perceived quality of male $j$ is $Q^{f}_{ij} = (-F_{p} \times k_{ij} + 1)^{-1}$. If the focal female has strategy of post-copulatory inbreeding preference ($F_{p}>0$), then the perceived quality of male $j$ is $Q^{f}_{ij} = F_{p} \times k_{ij} + 1$. A relative quality ($q^{f}_{ij}$) is then calculated for each male by dividing his $Q^{f}_{ij}$ by the sum of the $Q^{f}_{ij}$ values across all of a female's mates. These $q^{f}_{ij}$ values, which lie between zero and one, define the probability of paternity. As with mate choice, $q^{f}_{ij}$ values are placed into a vector whose elements sum to unity, and a random uniform number is used to sample paternity. Females produce $n$ offspring, so $q^{f}_{ij}$ values are sampled $n$ times independently and with replacement for each female to determine the realised distribution of sires.

@@

@@

@@

@@

However, when both pre-copulatory and post-copulatory inbreeding avoidance could evolve, their relative evolutionary dynamics depended on their relative costs. When one inbreeding strategy had zero cost, the other strategy initially evolved towards inbreeding avoidance (i.e., negative $M_{a}$ or $F_{a}$) but then evolved back towards random mating or fertilisation (i.e., $M_{a}=0$ or $F_{a}=0$) across generations following increasing evolution of inbreeding avoidance through the cost-free strategy (Figure 3B,D).

Evolution of post-copulatory inbreeding avoidance facilitated evolution of polyandry (Figure 3B,C), while monandry evolved in the absence of post-copulatory inbreeding avoidance (Figure 3A,D; polyandry was cost free in all these simulations).

@@

@@

@@

@@
